A multi-material HLLC Riemann solver with both elastic and plastic waves for 1D elastic-plastic flows

### 2019-04-27 Manuscript.tex
second

\subsection{State in region $\tilde{\mathbf{Q}}_{L(R)}$}\label{sec:tilde}

For cases (a,b,d,e) in Fig.\ref{fig:cases}, the material is totally yielding or totally not yielding, if we give a density $\rho_L^*$ in the region $\mathbf{Q}_L^*$, by the deductions in Section \ref{sec:rarefaction} and Section \ref{sec:shock}, all the unknowns in region $\mathbf{Q}_L^*$ can be solved out.

For cases (c,f), the meterial periods a yielding process from elastic to plastic. There is one more state $\tilde{\mathbf{Q}}_{L(R)}$ exists. In state $\tilde{\mathbf{Q}}_{L(R)}$, the derivative stress achieves the elastic limit 
\begin{equation}
  \tilde{s}_{xxL(R)} = \left\{ \begin{align}
	  &\frac{2}{3}Y_0 \quad  &\text{Case c},\\
	  &-\frac{2}{3}Y_0 \quad &\text{Case f},\\
	\end{align}
  \right.
\end{equation}
By (\ref{eq:rhosxx}), we can solve the density out as 
\begin{equation}
  \tilde{\rho}_{L(R)}= \left\{ \begin{align}
	  &\widetilde{\rho}_{L(R)} = \rho_{L(R)} \text{exp}\left(-\frac{Y_0}{2\mu}+\frac{3 s_{xxL(R)}}{4\mu}\right), \quad  & \text{Case c}, \\ 
	  &\widetilde{\rho}_{L(R)} = \rho_{L(R)} \text{exp}\left(\frac{Y_0}{2\mu}+\frac{3 s_{xxL(R)}}{4\mu}\right), \quad  & \text{Case f}. 
	\end{align}
  \right.
	\end{equation}
	As  the density in region $\mathbf{Q}^*_{L(R)}$ is known, all the pressure and velocity  can be solved by equations (\ref{eq:p_rho},\ref{eq:c_2rho},  \ref{eq:urho}) for a rarefaction wave  or by  equations (\ref{eq:shockp}, \ref{eq:shocku}) for a shock wave.

	\subsubsection{Elastic rarefaction wave in case c}
	For case c, the pressure within the elastic rarefaction wave is 
	\begin{equation}
	  p(\rho) = p_{L(R)}e^{\frac{\lambda_1}{\rho_{L(R)}}-\frac{\lambda_1}{\rho}} +e^{-\frac{\lambda_1}{\rho}}\int_{\rho_{L(R)}}^{\rho} f_2(x) e^{\frac{\lambda_1}{x}}dx, \quad  \tilde{\rho}_{L(R)} \le \rho \le \rho_{L(R)}.
	\end{equation}
The sonic speed is 
	\begin{equation}
	c(\rho) =  \sqrt{ a_0^2 \frac{\partial f}{\partial \eta} + \frac{p(\rho)}{\rho^2}\rho_0\Gamma_0 -\frac{\rho_0}{\rho^2}\Gamma_0 s_{xx} +\frac{4}{3}\frac{\mu}{\rho}} \quad   \tilde{\rho}_{L(R)} \le \rho \le \rho_{L(R)}.
	\end{equation}
With the sonic speed, the velocity can be solved as
\begin{equation}
  u(\rho) = \left\{ \begin{aligned}
	u_L - \int_{\rho_L}^\rho \frac{c(x)}{x} dx,  \quad   \tilde{\rho}_L \le \rho \le \rho_{L} \\
	u_R + \int_{\rho_R}^\rho \frac{c(x)}{x} dx,  \quad   \tilde{\rho}_R \le \rho \le \rho_{R} \\
	\end{aligned}
  \right.
\end{equation}
So the pressure and velocity after the elastic wave are 
\begin{equation}
  \tilde{p}_{L(R)} = p(\tilde{\rho}_{L(R)}), \quad \tilde{u}_{L(R)} = u(\tilde{\rho}_{L(R)}).
\end{equation}

\subsubsection{Elastic shock wave in  case f}
For a shock wave (case f ), the pressure is 
\begin{equation}
  \tilde{p}_{L(R)}= \frac{2t(c_1f(\tilde{\rho}_{L(R)}/\rho_0)+e_L)-(\sigma_{L(R)}+\tilde{s}_{xxL(R)})}{2tc_0-1},
\end{equation}
where $c_0 = \frac{1}{\rho_0\Gamma_0}$ and $c_1 = \frac{a_0^2}{\Gamma_0}$ and $ t=\frac{\rho_{L(R)} \tilde{\rho}_{L(R)}}{\tilde{\rho}_{L(R)}-\rho_{L(R)}}$. 
And the velocity is 
\begin{equation}
  \tilde{u}_L = u_L -\sqrt{\frac{\sigma_L-\tilde{\sigma}_L}{t}},
\end{equation}
for the left, and 
\begin{equation}
  \tilde{u}_R = u_R +\sqrt{\frac{\sigma_R-\tilde{\sigma}_R}{t}},
\end{equation}
for the right-going wave.
	\subsection{Functions of velocity and Cauchy stress in region $\mathbf{Q}^*_{L(R)}$}\label{sec:fufs}
	The deductions of  velocity and Cauchy stress  for rarefaction waves and shocks are given in Section \ref{sec:rarefaction} and Section \ref{sec:shock}. For convenient, we list them in this section in  different cases. 
	\subsubsection{ $\mathbf{Q}^*_{L(R)}$ in  rarefaction wave cases }
	For rarefaction waves  in region
	The deviatoric stress is 
	\begin{equation}
	  s_{xxL(R)}^*= \left\{ \begin{aligned}
		  & s_{xxL(R)} \quad &\text{case (b, e)}, \\
		  & -\frac{4}{3}\mu\text{ln}\left(\frac{\rho_{L(R)}^*}{\rho_{L(R)}}\right)+s_{xxL(R)} \quad  &\text{case (a,d)},\\
		  & \frac{2}{3}Y_0 \quad  &\text{case (c)},\\
		  & -\frac{2}{3}Y_0 \quad &\text{case (f)}.\\
		\end{aligned}
	  \right.
	\end{equation}
The  pressure for case (a,b,c) is 
\begin{equation}
  p_{L(R)}^* = \left\{\begin{aligned}
	  &	p_{L(R)}e^{\frac{\lambda_1}{\rho_{L(R)}}-\frac{\lambda_1}{\rho_{L(R)}^*}} +e^{-\frac{\lambda_1}{\rho_{L(R)}^*}}\int_{\rho_{L(R)}}^{\rho_{L(R)}^*} f_2(x) e^{\frac{\lambda_1}{x}}dx. \quad & \text{case (a,b)},\\
	  &	\tilde{p}_{L(R)}e^{\frac{\lambda_1}{\tilde{\rho}_{L(R)}}-\frac{\lambda_1}{\rho_{L(R)}^*}} +e^{-\frac{\lambda_1}{\rho_{L(R)}^*}}\int_{\tilde{\rho}_{L(R)}}^{\rho_{L(R)}^*} f_2(x) e^{\frac{\lambda_1}{x}}dx. \quad & \text{case (c)},
  \end{aligned}
\right.
\end{equation} 
after get the pressure we can give the sonic speed as
\begin{equation}
  c_{L(R)}^{*2} = \left\{ \begin{aligned}
	&  a_0^2 \frac{\partial f}{\partial \eta} + \frac{p_{L(R)}^*}{\rho_{L(R)}^{*2}}\rho_0\Gamma_0 -\frac{\rho_0}{\rho_{L(R)}^{*2}}\Gamma_0 s_{xx} +\frac{4}{3}\frac{\mu}{\rho_{L(R)}^*} \quad & \text{case a},\\
	&	a_0^2 \frac{\partial f}{\partial \eta} + \frac{p}{\rho_{L(R)}^{*2}}\rho_0\Gamma_0 -\frac{\rho_0}{\rho_{L(R)}^{*2}}\Gamma_0 s_{xxL(R)}^*  \quad  & \text{case (b,c)}.
	\end{aligned}
  \right.
\end{equation}
then the velocity can be solved as
\begin{equation}
  u_{L}^* =\left\{ \begin{aligned} 
	  &u_L - \int_{\rho_L}^{\rho_L^*} \frac{c}{\rho} d\rho \quad \text{case a,b}, \\
	  &\tilde{u}_L - \int_{\tilde{\rho}_L}^{\rho_L^*} \frac{c}{\rho} d\rho \quad \text{case c}, \\
	\end{aligned}
  \right.
 u_{R}^* =\left\{ \begin{aligned} 
	 &u_R + \int_{\rho_R}^{\rho_R^*} \frac{c}{\rho} d\rho \quad \text{case a,b}, \\
	 &\tilde{u}_R + \int_{\tilde{\rho}_R}^{\rho_R^*} \frac{c}{\rho} d\rho \quad \text{case c}. \\
	\end{aligned}
  \right.
\end{equation}

>>>>

\subsection{States for one wave cases (a,b, d,e)}
For cases (a,b,d,e) in Fig.\ref{fig:cases}, the material is totally yielding or totally not yielding, if we give a density $\rho_L^*$ in the region $\mathbf{Q}_L^*$, by the deductions in Section \ref{sec:rarefaction} and Section \ref{sec:shock}, all the unknowns in region $\mathbf{Q}_L^*$ can be solved out.

\subsubsection{Rarefaction  wave cases (a,b)}
For Rarefaction wave, we not only  need to solve the state after the wave in region $\mathbf{Q}_{L(R)}^*$, but also  need to know the states inside the expansion region.

First we give the function of $s_{xx}$,
\begin{equation}
  s_{xx}(\rho) = \left\{\begin{aligned}
	& -\frac{4}{3}\mu\text{ln}\left(\frac{\rho}{\rho_{L(R)}}\right)+s_{xxL(R)} \quad  &\text{case (a)},\\
	& \frac{2}{3}Y_0 &\text{case (b)}.
  \end{aligned} \right.
  \end{equation}
Then we give the pressure,
\begin{equation}
  p(\rho)=p_{L(R)}e^{\frac{\lambda_1}{\rho_{L(R)}}-\frac{\lambda_1}{\rho}} +e^{-\frac{\lambda_1}{\rho}}\int_{\rho_{L(R)}}^{\rho} f_2(x) e^{\frac{\lambda_1}{x}}dx. \quad & \text{case (a,b)}.
\end{equation}
and sonic speed,
\begin{equation}
  c(\rho) = \left\{ \begin{aligned}
	&  \sqrt{a_0^2 \frac{\partial f}{\partial \eta} + \frac{p(\rho)}{\rho^2}\rho_0\Gamma_0 -\frac{\rho_0}{\rho^2}\Gamma_0 s_{xx}(\rho) +\frac{4}{3}\frac{\mu}{\rho}} \quad & \text{case (a)},\\
	&	\sqrt{a_0^2 \frac{\partial f}{\partial \eta} + \frac{p(\rho)}{\rho^2}\rho_0\Gamma_0 -\frac{\rho_0}{\rho^2}\Gamma_0 s_{xx}(\rho)}  \quad  & \text{case (b)}.
	\end{aligned}\right.
\end{equation}
Then we can get  the  function of  velocity
\begin{equation}
  u(\rho) =\left\{ \begin{aligned} 
	&u_L - \int_{\rho_L}^{\rho} \frac{c}{\rho} d\rho \quad \text{case (a,b) Left} , \\
	 &u_R + \int_{\rho_R}^{\rho} \frac{c}{\rho} d\rho \quad   \rho_R^* \le \rho \le \rho_R \quad \text{case (a,b) Right}, \\
	\end{aligned}
  \right.
\end{equation}

\subsubsection{Shock wave cases (d,e)}
For shock waves, the function of deviatoric is given as 
\begin{equation}
  s_{xx}(\rho) = \left\{\begin{aligned}
	& \begin{aligned}
	  & s_{xxL(R)} \quad  &\rho = \rho_{L(R)}\\
	& -\frac{4}{3}\mu\text{ln}\left(\frac{\rho}{\rho_{L(R)}}\right)+s_{xxL(R)}\quad &  \rho =\rho_{L(R)}^* \end{aligned}
	\quad  &\text{case (d)},\\
	& \frac{2}{3}Y_0 &\text{case (e)}.
  \end{aligned} \right.
  \end{equation}
And the pressure is given as
\begin{equation}
  p(\rho)= \left\{\begin{aligned} 
	& p_{L(R)}   \quad  & \rho = \rho_{L(R)} \\
	& \frac{2t(c_1f(\rho/\rho_0)+e_L)-(\sigma_{L(R)}+s_{xx}(\rho))}{2tc_0-1} \quad  &  \rho=\rho_{L(R)}^* 
  \end{aligned}\right.
  \text{ case (d,e)},
\end{equation}
where $c_0 = \frac{1}{\rho_0\Gamma_0}$ and $c_1 = \frac{a_0^2}{\Gamma_0}$ and $ t=\frac{\rho_{L(R)} \rho_{L(R)}^*}{\rho_{L(R)}^*-\rho_{L(R)}}$. 
The velocity is given as
\begin{equation}
  u(\rho) = \left\{ \begin{aligned} u_{L(R)} \quad  &\rho =\rho_{L(R)}\\
	u_L -\sqrt{\frac{\sigma_L-\sigma(\rho)}{t}} \quad &\rho = \rho_L^*\\ 
	u_R +\sqrt{\frac{\sigma_R-\sigma(\rho)}{t}} \quad & \rho = \rho_R^*\\
  \end{aligned}
  \right.
\end{equation}

\subsection{States for two wave cases (c,f)}

For cases (c,f), the meterial periods a yielding process from elastic to plastic. There are two waves and  one more state $\tilde{\mathbf{Q}}_{L(R)}$ exists. In state $\tilde{\mathbf{Q}}_{L(R)}$, the derivative stress achieves the elastic limit. 
\begin{equation}
 \tilde{s}_{xxL(R)} = \left\{ \begin{align}
	  &\frac{2}{3}Y_0 \quad  &\text{Case c},\\
	  &-\frac{2}{3}Y_0 \quad &\text{Case f},\\
	\end{align}
  \right.
\end{equation}
By (\ref{eq:rhosxx}),at state $\tilde{\mathbf{Q}}_{L(R)}$ we can solve the density out as 
\begin{equation}
  \tilde{\rho}_{L(R)}= \left\{ \begin{align}
	  &\widetilde{\rho}_{L(R)} = \rho_{L(R)} \text{exp}\left(-\frac{Y_0}{2\mu}+\frac{3 s_{xxL(R)}}{4\mu}\right), \quad  & \text{Case c}, \\ 
	  &\widetilde{\rho}_{L(R)} = \rho_{L(R)} \text{exp}\left(\frac{Y_0}{2\mu}+\frac{3 s_{xxL(R)}}{4\mu}\right), \quad  & \text{Case f}. 
	\end{align}
  \right.
	\end{equation}
\subsubsection{Rarefaction wave case (c)}
For rarefaction wave case,
we give the function of $s_{xx}$ at first,
\begin{equation}
  s_{xx}(\rho) = \left\{\begin{aligned}
	& -\frac{4}{3}\mu\text{ln}\left(\frac{\rho}{\rho_{L(R)}}\right)+s_{xxL(R)} \quad  & \rho_{L(R)} \ge \rho \ge \tilde{\rho}_{L(R)}  ,\\
	& \frac{2}{3}Y_0 & \tilde{\rho}_{L(R)} \ge \rho \ge \rho_{L(R)}^* .
  \end{aligned} \right. \quad \text{case (c)}.
  \end{equation}
And sonic speed,
\begin{equation}
  c(\rho) = \left\{ \begin{aligned}
	&  \sqrt{a_0^2 \frac{\partial f}{\partial \eta} + \frac{p(\rho)}{\rho^2}\rho_0\Gamma_0 -\frac{\rho_0}{\rho^2}\Gamma_0 s_{xx}(\rho) +\frac{4}{3}\frac{\mu}{\rho}} \quad  & \rho_{L(R)} \ge \rho \ge \tilde{\rho}_{L(R)}\\
	&	\sqrt{a_0^2 \frac{\partial f}{\partial \eta} + \frac{p(\rho)}{\rho^2}\rho_0\Gamma_0 -\frac{\rho_0}{\rho^2}\Gamma_0 s_{xx}(\rho)}  \quad  &   \tilde{\rho}_{L(R)} \ge \rho \ge \rho_{L(R)}^*.
  \end{aligned}\right. \quad \text{case (c)}.
\end{equation}
Then we can get  the   function of  velocity
\begin{equation}
  u(\rho) =\left\{ \begin{aligned} 
	&u_L - \int_{\rho_L}^{\rho} \frac{c(x)}{x} dx, \quad &  \rho_L\ge \rho\ge \rho_L^* , \\
	&u_R + \int_{\rho_R}^{\rho} \frac{c(x)}{x} dx, \quad &  \rho_L\ge \rho\ge  \rho_R^* , 
	\end{aligned}
  \right. \quad \text{case (c)}.
\end{equation}
\subsubsection{Shock wave case (f)}
For shock waves the function of deviatoric is given as 
\begin{equation}
  s_{xx}(\rho) = \left\{\begin{aligned}
  & -\frac{4}{3}\mu\text{ln}\left(\frac{\rho}{\rho_{L(R)}}\right)+s_{xxL(R)}\quad &  \rho =\rho_{L(R)}, \tilde{\rho}_{L(R)},\\
	& \frac{2}{3}Y_0 &\rho = \rho_{L(R)}^*,
  \end{aligned} \right. \quad  \text{case (f)}.
  \end{equation}

And the pressure is given as
\begin{equation}
  p(\rho)= \left\{\begin{aligned} 
	& p_{L(R)}   \quad  & \rho = \rho_{L(R)} \\
	& \frac{2t(c_1f(\rho/\rho_0)+e_L)-(\sigma_{L(R)}+s_{xx}(\rho))}{2t_1c_0-1} \quad  &  \rho=\tilde{\rho}_{L(R)}\\ 
	& \frac{2t(c_1f(\rho/\rho_0)+\tilde{e}_L)-(\tilde{\sigma}_{L(R)}+s_{xx}(\rho))}{2t_2c_0-1} \quad  &  \rho=\rho_{L(R)}^*\\ 
  \end{aligned}\right.
  \text{ case (f)},
\end{equation}
where $c_0 = \frac{1}{\rho_0\Gamma_0}$ and $c_1 = \frac{a_0^2}{\Gamma_0}$ and $ t_2=\frac{\rho_{L(R)} \tilde{\rho}_{L(R)}}{\tilde{\rho}_{L(R)}-\rho_{L(R)}}$, $ t_1=\frac{\tilde{\rho}_{L(R)} \rho_{L(R)}^*}{\rho_{L(R)}^*-\tilde{\rho}_{L(R)}}$

The velocity is given as
\begin{equation}
  u(\rho) = \left\{ \begin{array}{ll}
	u_{L(R)} \quad  & \rho =\rho_{L(R)}\\
		u_L -\sqrt{\frac{\sigma_L-\sigma(\rho)}{t_1}} \quad &\rho = \tilde{\rho}_L\\ 
		\tilde{u}_L -\sqrt{\frac{\tilde{\sigma}_L-\sigma(\rho)}{t_2}} \quad &\rho = \rho_L^*\\ 
		u_R +\sqrt{\frac{\sigma_R-\sigma(\rho)}{t_1}} \quad &\rho = \tilde{\rho}_R\\ 
  \tilde{u}_R +\sqrt{\frac{\tilde{\sigma}_R-\sigma(\rho)}{t_2}} \quad &\rho = \rho_R^*\\ \end{array}
  \right. \quad \text{case (f)}.
\end{equation}
